---
title: "code"
author: "Ye Yao"
date: "2022-12-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(survival)
library(survminer)
library(gtsummary)
library(dplyr)
library(plyr)
library(tableone)  
library(kableExtra)
library(randomForestSRC)
library(pec)
library(mice)
library(corrplot)
library(circlize)
library(ggplot2)
library(tidyverse)
library(RColorBrewer)
library(survival)
library(survminer)
library(mgcv)
library(lattice)
library(MASS)
library(nnet)
library(mice)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(AUC)
library(tableone)  
library(survival)
library(kableExtra)
library(forestplot)
library(stringr)
library(rms)
library(pec)
library(ipred)
```
```{r}
# read data
rawdata <- read.csv("rawraw1.csv",sep=" ")

#data preparation
#choose year
raw1 <- rawdata[rawdata$Year.of.diagnosis<=2015&rawdata$Year.of.diagnosis>=2010,]

#rename colname
colnames(raw1) <- c("age1","sex","year","diagnose","ICDO31","ICDO32","race","age2","months","flag","ER","PR","HER","tumor_size","stage1","RX1","reginal")
```
#factor age
#raw1$age <- factor(raw1$age,levels = c("01-04 years", "05-09 years", "10-14 years","15-19 years","20-24 years","25-29 year","30-34 years","35-39 years","40-44 years","45-49 years","50-54 years","55-59 years","60-64 years","65-69 years","70-74 years","75-79 years","80-84 years","85+ years"))
#raw1$age <- factor(raw1$age, labels=c("01-04", "05-09", "10-14","15-19","20-24","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80-84","85+"))

#deal with age
```{r}
raw1$age <- as.numeric(lapply(raw1$age2,function(x) substr(x,1,2)))
```

#deal with race
```{r}
r1 <- c("Chinese","Japanese","Filipino","Korean (1988+)","Vietnamese (1988+)","Laotian (1988+)","Hmong (1988+)","Kampuchean (1988+)","Thai (1994+)","Asian Indian (2010+)","Asian Indian or Pakistani, NOS (1988+)")
r2 <- c("White")
r3 <- c("Black")
r4 <- c("American Indian/Alaska Native")
r5 <- c("Hawaiian","Micronesian, NOS (1991+)","Chamorran (1991+)","Guamanian, NOS (1991+)","Polynesian, NOS (1991+)","Tahitian (1991+)","Samoan (1991+)","Tongan (1991+)","Melanesian, NOS (1991+)","Fiji Islander (1991+)","New Guinean (1991+)","Pacific Islander, NOS (1991+)")

raw1$race_6 <- 0
for (i in 1:dim(raw1)[1]){
  if (raw1$race[i] %in% r1){
    raw1$race_6[i] <- 0 #"Asian"
  }
  else if (raw1$race[i] %in% r2){
    raw1$race_6[i] <- 1 #"White"
  }
  else if (raw1$race[i] %in% r3){
    raw1$race_6[i] <- 2 #"Black"
  }
  else if (raw1$race[i] %in% r4){
    raw1$race_6[i] <- 3 #"America Indian"
  }
  else if (raw1$race[i] %in% r5){
    raw1$race_6[i] <- 4 #"Native Hawaii or other Pacific Islander"
  }
  else{
    raw1$race_6[i] <- 5 #"Others"
  }
}
```
#deal with stage
```{r}
raw1$stage <- NA
raw1$stage[raw1$stage1=="0"] <- 0
raw1$stage[raw1$stage1=="I"] <- 1
raw1$stage[raw1$stage1 %in% c("IIA","IIB")] <- 2
raw1$stage[raw1$stage1 %in% c("IIIA","IIIB","IIIC","IIINOS")] <- 3
raw1$stage[raw1$stage1=="IV"] <- 4
```

#deal with ER status
```{r}
raw1$ER[raw1$ER %in% c("Borderline/Unknown","Recode not available")] <- NA
raw1$ER[raw1$ER=="Positive"] <- 1
raw1$ER[raw1$ER=="Negative"] <- 0
```
#deal with PR status
```{r}
raw1$PR[raw1$PR %in% c("Borderline/Unknown","Recode not available")] <- NA
raw1$PR[raw1$PR=="Positive"] <- 1
raw1$PR[raw1$PR=="Negative"] <- 0
```
#deal with HER status
```{r}
raw1$HER[raw1$HER %in% c("Borderline/Unknown","Recode not available")] <- NA
raw1$HER[raw1$HER=="Positive"] <- 0
raw1$HER[raw1$HER=="Negative"] <- 1
```
#construct new variable
```{r}
raw1$ER_PR <- NA
raw1$ER_PR[raw1$ER==1|raw1$PR==1] <- 1
raw1$ER_PR[raw1$ER==1&raw1$PR==1] <- 2
raw1$ER_PR[raw1$ER==0&raw1$PR==0] <- 0
```
#deal with RX1
```{r}
raw1 <- raw1[raw1$RX1!=99,]
raw1$surgery <- NA
raw1$surgery[raw1$RX1>0] <- 1
raw1$surgery[raw1$RX1==0] <- 0

```
#deal with tumor size
```{r}
raw1$tumor_size <- as.integer(raw1$tumor_size)
raw1$real_tumor <- NA
for (i in 1:dim(raw1)[1]){
  if (raw1$tumor_size[i] %in% c(990,995,996,997,998,999)){
    raw1$real_tumor[i] <- NA
  }
  else if (raw1$tumor_size[i]<=100 | raw1$tumor_size[i] == 991){
    raw1$real_tumor[i] <- 10 
  }
  else if (raw1$tumor_size[i]<=200 | raw1$tumor_size[i] == 992){
    raw1$real_tumor[i] <- 20
  }
  else if (raw1$tumor_size[i]<=300 | raw1$tumor_size[i] == 993){
    raw1$real_tumor[i] <- 30
  }
  else if (raw1$tumor_size[i]<=400 | raw1$tumor_size[i] == 994){
    raw1$real_tumor[i] <- 40 
  }
  else if (raw1$tumor_size[i]<=500){
    raw1$real_tumor[i] <- 50
  }
  else if (raw1$tumor_size[i]>500){
    raw1$real_tumor[i] <- 60
  }
}
```
#deal with survival month
```{r}
raw1$months <- as.integer(raw1$months)
raw1 <- raw1[raw1$months>=0,]
raw1 <- raw1[!is.na(raw1$flag),]
```
#deal with flag
```{r}
raw1$status  <- NA
for (i in 1:dim(raw1)[1]){
  if (raw1$flag[i]=="Complete dates are available and there are 0 days of survival"){
    raw1$status[i] <- 2 
  }
  else if (raw1$flag[i]=="Complete dates are available and there are more than 0 days of survival"){
    raw1$status[i] <- 2 
  }
  else if (raw1$flag[i]=="Incomplete dates are available and there cannot be zero days of follow-up"){
    raw1$status[i] <- 1
  }
  else if (raw1$flag[i]=="Incomplete dates are available and there could be zero days of follow-up"){
    raw1$status[i] <- 1 
  }
}
```
#deal with bad tumor number #delete this variable
#raw1$tumor_number_bad[raw1$tumor_number_bad=="Unknown"] <- NA
#raw1$tumor_number_bad <- as.integer(raw1$tumor_number_bad)

#deal with good tumor number #delete this variable
#raw1$tumor_number_good <- as.integer(raw1$tumor_number_good)

#deal with sex
```{r}
raw1$sex[raw1$sex=="Female"] <- 1
raw1$sex[raw1$sex=="Male"] <- 0
raw1$sex <- as.numeric(raw1$sex)

#deal with reginal
raw1$reginal <- as.numeric(raw1$reginal)
raw1$reginal[raw1$reginal>90] <- NA

data <- raw1 %>% select("age","sex","year","months","status","HER","race_6","stage","ER_PR","surgery","real_tumor","reginal")

#factorize
data$HER <- factor(data$HER)
data$race_6 <- factor(data$race_6)
data$stage <- factor(data$stage)
data$ER_PR <- factor(data$ER_PR)
data$surgery <- factor(data$surgery)
data$sex <- factor(data$sex)
data$real_tumor <- factor(data$real_tumor)

#five year status
data$status5 <- NA
data$status5[data$months>60] <- 0
data$status5[data$months<=60&data$status==1] <- 1
data$status5[data$months<=60&data$status==2] <- 2

train <- data[data$year<=2014,]
test <- data[data$year==2015,]
test <- na.omit(test)

# complete data
complete_train <- na.omit(train)
```

#Missing value analysis
```{r}
table1_data <- data
table1_data$Incomplete = rep(0,nrow(table1_data))
for(i in 1:nrow(table1_data)){
  if(sum(is.na(table1_data[i,]))!=0){table1_data$Incomplete[i]=1}
}
## Table 1 generation (outcome is generated using original data with 1707 samples)
table1_data %>%
  select("age","sex","HER","race_6","stage","ER_PR","surgery","real_tumor","reginal","Incomplete")%>%
  tbl_summary(by = Incomplete,
              missing = "no",
              type = list(age ~ 'continuous',
                          real_tumor ~ 'continuous'),
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(2,2),
                            all_categorical() ~ c(0,2))
  ) %>%
  add_n %>%
  add_p(test = list(all_continuous() ~ "t.test",
                    all_categorical() ~ "chisq.test"),
        test.args = all_tests("t.test") ~ list(var.equal = TRUE),## Important argument! 
        pvalue_fun = function(x) style_pvalue(x, digits = 2)) %>%
  bold_p(t = 0.05) %>%
  bold_labels

imp <- mice(train, m=5,printFlag = T,seed=10000)
trainc <- complete(imp)
write.csv(x = trainc,file = "train.csv")

trainc$status5 <- NA
trainc$status5[trainc$months>60] <- 0
trainc$status5[trainc$months<=60&trainc$status==1] <- 1
trainc$status5[trainc$months<=60&trainc$status==2] <- 2

trainc$months[trainc$months>60] <- 61
```
#Cox regression
```{r}
time <- "months"
status <- "status5"
y<- Surv(trainc[,time],trainc[,status])
```
#trainc$reginal_f <- NA
#trainc$reginal_f[trainc$reginal>4] <- 1
#trainc$reginal_f[trainc$reginal<=4] <- 0
#trainc$reginal_f <- factor(trainc$reginal_f)
```{r}
fit_cox1 <- coxph(Surv(months,status5==2)~sex+age+ER_PR+race_6+real_tumor+surgery+HER+stage+reginal+reginal:real_tumor,data=trainc)
#fit_cox1 <- coxph(Surv(months,status5==2)~sex+age+ER_PR+race_6+factor(real_tumor)+surgery+HER+stage+reginal_f:factor(real_tumor),data=trainc)
summary(fit_cox1)

result <- read.csv("forest.csv",header=TRUE,sep=',')#读入数据集
#knitr::kable(head(data),booktabs = TRUE,align='c')
head(result)
```
```{r}
fig1<- forestplot(result[,c(1,2,3)], #告诉函数，合成的表格result的第1，5，6列还是显示数字
                  mean=result[,4],   #告诉函数，表格第2列为HR，它要变成森林图的小方块
                  lower=result[,5],  #告诉函数表格第3列为5%CI，
                  upper=result[,6],  #表格第5列为95%CI，它俩要化作线段，穿过方块
                  zero=1,            #告诉函数，零线或参考线为HR=1即x轴的垂直线
                  boxsize=0.6,      
                  graph.pos= 2,
                  hrzl_lines=list("1" = gpar(lty=1,lwd=2),
                                  "2" = gpar(lty=2)),
                  xticks=c(0,0.8,1.6,2.4,3.2,4,4.8,5.6,6.4) ,
                  is.summary=c(T,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F,F),
                  txt_gp=fpTxtGp(label=gpar(cex=1.25),
                                 ticks=gpar(cex=1.1),
                                 xlab=gpar(cex = 1.2),
                                 title=gpar(cex = 1.2)),
                  graphwidth=unit(150,"mm"),
                  lwd.zero=1,
                  lwd.ci=1.5,
                  lwd.xaxis=2, 
                  lty.ci=1.5,
                  ci.vertices =T,
                  ci.vertices.height=0.2, 
                  clip=c(0.1,6.4),
                  ineheight=unit(4, 'mm'), 
                  line.margin=unit(4, 'mm'),
                  colgap=unit(10, 'mm'),
                  fn.ci_norm="fpDrawDiamondCI",
                  col=fpColors(box ='#021eaa', 
                               lines ='#021eaa', 
                               zero = "black"))       #森林图应插在图形第2列
fig1
```
```{r}
test <- test %>% select("age","sex","months","HER","race_6","stage","ER_PR","surgery","real_tumor","reginal","status5")
test$status5 <- 0
test$status5[test$months>48] <- 0
test$status5[test$months<=48&test$status==1] <- 1
test$status5[test$months<=48&test$status==2] <- 2

test$months[test$months>60] <- 61

fp <- predict(fit_cox1) 
summary(fit_cox1)$concordance 
library(Hmisc) 
options(scipen=200) 
with(trainc,rcorr.cens(fp,Surv(months, status5==1)))

pre1<-predict(fit_cox1, newdata=test)
rcorrcens(Surv(months,status5==2) ~pre1, data = test)

coxph_err <- sapply(1:500,function(b){
  if(b%%50==0)
    cat("cox boostrap:",b,"\n")
    train <- sample(1:nrow(trainc),nrow(trainc),replace=T)
    coxph_13.obj <- tryCatch({coxph(Surv(months,status5==2)~sex+age+ER_PR+race_6+real_tumor+surgery+HER+stage+reginal+reginal:real_tumor,data=trainc[train,])},error=function(ex){NULL})
  
  if(!is.null(coxph_13.obj)){
    get.cindex(trainc$months[-train],trainc$status5[-train],predict(coxph_13.obj,trainc[-train,]))
  }
  else NA
})

mean(coxph_err,na.rm=T)
```
#nomogram
```{r}

train_cox <- trainc %>% dplyr::select("age","sex","months","HER","race_6","stage","ER_PR","surgery","real_tumor","reginal","status5")
colnames(train_cox) <- c("Age","Sex","months","HER","Race","Stage","ER_PR","Surgery","Tumor_Size","Reginal","status5")
ddist <- datadist(train_cox) 
options(datadist='ddist')
f <- cph(Surv(months,status5==2)~Age+Sex+HER+Race+Stage+ER_PR+Surgery+Tumor_Size+Reginal,x=T, y=T,surv=T,data=train_cox, time.inc=60)
surv <- Survival(f)
nom2 <- nomogram(f,fun=function(x) surv(60, x),lp=F, funlabel="5 year survival",maxscale=100,fun.at=c(0.99, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3,0.2,0.1,0.05))   #构建Nomogram  
plot(nom2)   #输出Nomogram图


```


#Overall survival
```{r}
fit.1 <- survfit(Surv(months,status5==2)~1, data=trainc)
s.1 <- ggsurvplot(fit.1,data=trainc,conf.int = TRUE,conf.int.alpha=0.5,
                  conf.int.style="step",
                  #color="strata",
                  palette="#e63946", 
                  #legend.labs="Censor", 
                  #legend.title="",
                  #legend = c(0.15, 0.5),
                  #font.legend = 18,
                  surv.median.line = "hv",
                  risk.table = 'nrisk_cumcensor',tables.height = 0.2,
                  #xlim = c(0,990),
                  break.y.by = 0.2,
                  tables.theme = theme_cleantable(),
                  #ggtheme = theme_bw(),
                  risk.table.y.text.col = T,# colour risk table text annotations.
                  risk.table.height = 0.15, # the height of the risk table
                  risk.table.y.text = FALSE,# show bars instead of names in text annotations
                  #title="30days survival",
                  ylab="Probability of survival",xlab = "Days since diagnosis")
```
#更改横纵坐标
#xlim = c(0,990),
s.1
```{r}
f1 <- survfit(Surv(months,status5==2)~trainc$stage, data=trainc)
s1 <- ggsurvplot(f1,data=trainc, pval = F,#conf.int = TRUE,
                 conf.int.alpha=0.3,#conf.int.style="step",
                 pval.coord = c(5, 0.05), #调节Pval的位置
                 pval.method = T,
                 pval.method.coord = c(0, 0.05),
                 surv.median.line = "hv", #添加中位生存曲线
                 #risk.table = 'nrisk_cumcensor',#tables.height = 0.2,
                 #surv.scale = 'percent',
                 break.y.by=0.2,
                 risk.table.y.text = FALSE,
                 #palette=c("#e63946", "#457b9d","#81b29a"), 
                 #legend.labs=c("Normal (3.11-5.18mmol/L)","Low (≤3.11mmol/L)","High (≥5.18mmol/L)"), #标签
                 legend.title="Stage",
                 #legend = c(0.28, 0.45),
                 #font.legend = 12,
                 #title="Survival Group By Stage",
                 ylab="Probability of survival",xlab = "Months since diagnosis", #更改横纵坐标
                 break.x.by = 10)
s1

f2 <- survfit(Surv(months,status5==2)~trainc$ER_PR, data=trainc)
s2 <- ggsurvplot(f2,data=trainc, pval = T,#conf.int = TRUE,
                 conf.int.alpha=0.3,#conf.int.style="step",
                 #pval.coord = c(5, 0.05), #调节Pval的位置
                 pval.method = T,
                 #pval.method.coord = c(0, 0.05),
                 #surv.median.line = "hv", #添加中位生存曲线
                 #risk.table = 'nrisk_cumcensor',#tables.height = 0.2,
                 #surv.scale = 'percent',
                 break.y.by=0.2,
                 #risk.table.y.text = FALSE,
                 #palette=c("#e63946", "#457b9d","#81b29a"), 
                 #legend.labs=c("Normal (3.11-5.18mmol/L)","Low (≤3.11mmol/L)","High (≥5.18mmol/L)"), #标签
                 legend.title="ER_PR",
                 #legend = c(0.28, 0.45),
                 font.legend = 12,
                 #title="Survival Group By Cholesterol",
                 ylab="Probability of survival",xlab = "Months since diagnosis", #更改横纵坐标
                 break.x.by = 10)
s2

f3 <- survfit(Surv(months,status5==2)~trainc$HER, data=trainc)
s3 <- ggsurvplot(f3,data=trainc, pval = T,#conf.int = TRUE,
                 conf.int.alpha=0.3,#conf.int.style="step",
                 #pval.coord = c(5, 0.05), #调节Pval的位置
                 pval.method = T,
                 #pval.method.coord = c(0, 0.05),
                 #surv.median.line = "hv", #添加中位生存曲线
                 #risk.table = 'nrisk_cumcensor',#tables.height = 0.2,
                 #surv.scale = 'percent',
                 break.y.by=0.05,
                 #risk.table.y.text = FALSE,
                 #palette=c("#e63946", "#457b9d","#81b29a"), 
                 #legend.labs=c("Normal (3.11-5.18mmol/L)","Low (≤3.11mmol/L)","High (≥5.18mmol/L)"), #标签
                 legend.title="HER",
                 #legend = c(0.28, 0.45),
                 font.legend = 12,
                 #ylim=c(0.6,1),
                 #title="Survival Group By Cholesterol",
                 ylab="Probability of survival",xlab = "Months since diagnosis", #更改横纵坐标
                 break.x.by = 10)
s3

f4 <- survfit(y~complete_data$PR, data=complete_data)
s4 <- ggsurvplot(f4,data=complete_data, pval = F,#conf.int = TRUE,
                 conf.int.alpha=0.3,#conf.int.style="step",
                 pval.coord = c(5, 0.05), #调节Pval的位置
                 pval.method = T,
                 pval.method.coord = c(0, 0.05),
                 surv.median.line = "hv", #添加中位生存曲线
                 #risk.table = 'nrisk_cumcensor',#tables.height = 0.2,
                 #surv.scale = 'percent',
                 break.y.by=0.2,
                 risk.table.y.text = FALSE,
                 #palette=c("#e63946", "#457b9d","#81b29a"), 
                 #legend.labs=c("Normal (3.11-5.18mmol/L)","Low (≤3.11mmol/L)","High (≥5.18mmol/L)"), #标签
                 legend.title="PR",
                 #legend = c(0.28, 0.45),
                 font.legend = 12,
                 #title="Survival Group By Cholesterol",
                 ylab="Probability of survival",xlab = "Months since diagnosis", #更改横纵坐标
                 break.x.by = 10)

f5 <- survfit(Surv(months,status5==2)~complete_data$race1, data=complete_data)
s5 <- ggsurvplot(f5,data=complete_data, pval = F,#conf.int = TRUE,
                 conf.int.alpha=0.3,#conf.int.style="step",
                 #pval.coord = c(5, 0.05), #调节Pval的位置
                 #pval.method = T,
                 #pval.method.coord = c(0, 0.05),
                 surv.median.line = "hv", #添加中位生存曲线
                 #risk.table = 'nrisk_cumcensor',#tables.height = 0.2,
                 #surv.scale = 'percent',
                 break.y.by=0.2,
                 risk.table.y.text = FALSE,
                 #palette=c("#e63946", "#457b9d","#81b29a"), 
                 #legend.labs=c("Normal (3.11-5.18mmol/L)","Low (≤3.11mmol/L)","High (≥5.18mmol/L)"), #标签
                 legend.title="Race",
                 #legend = c(0.28, 0.45),
                 font.legend = 12,
                 #title="Survival Group By Cholesterol",
                 ylab="Probability of survival",xlab = "Months since diagnosis", #更改横纵坐标
                 break.x.by = 10)
s5


```
#random survival forest
```{r}
surv_rf <- as.formula(Surv(months,status5==2)~sex+age+ER_PR+race_6+factor(real_tumor)+surgery+HER+stage+reginal)
o13 <- tune(surv_rf,trainc,ntreeTry = 50)

o <-  rfsrc(Surv(months,status5==2)~sex+age+ER_PR+race_6+real_tumor+surgery+HER+stage+reginal, trainc, nsplit = 10, ntree = 50,block.size = 1)
print(o)

get.cindex(time = trainc$months, censoring = trainc$status5==1, predicted = o$predicted.oob[,2])

pred <- predict(o, test, OOB=TRUE,na.action="na.impute")
#pred <- predict(o, newdata=test)
print(o)
print(pred)

get.cindex(time = test$months, censoring = test$status5==1, predicted = pred$predicted.oob[,2])

prederror <- pec(o,data=trainc,formula=Hist(months,status5)~sex+age+ER_PR+race_6+surgery+HER+stage+reginal,splitMethod = "bootcv",B=50)
```